---
title: "Editing a research paper with a review gate"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Editing a research paper with a review gate}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(superdocsr)
```

## The problem this solves

An AI that edits your manuscript is useful right up to the moment it changes
something you did not want changed. In a paper, that is not a cosmetic
problem: a rewritten hedge becomes an overclaim, a "tightened" result becomes
a different result, and neither is obvious three drafts later.

So this package does not apply edits. It proposes them, shows you what they
are, and applies exactly the ones you name. The workflow is five calls:

    sd_upload()  ->  sd_edit()  ->  sd_wait()  ->  sd_changes()  ->  sd_approve()  ->  sd_export()

and the gate sits in the middle of it, not as an option you can switch on.

## Running this vignette

Everything below actually runs, with no API key and no network. The client's
HTTP transport is a plain function, so this vignette supplies one that returns
recorded SuperDocs payloads. That is the same seam the package's test suite
uses, and it is available to you: if you are building on top of `superdocsr`,
you can test your own code the same way.

```{r fake-api}
# Recorded payloads, shaped exactly like the documented API responses.
payloads <- list(
  upload = list(
    html = "<div data-chunk-id=\"c1\"><h1>Effects of X on Y</h1></div>",
    session_id = "paper-2026", filename = "manuscript.docx",
    chunks_count = 96, version_id = "v_1",
    page_setup = list(width_in = 8.27, height_in = 11.69, orientation = "portrait")
  ),
  started = list(job_id = "job_7f3a", session_id = "paper-2026", status = "pending"),
  review = list(
    job_id = "job_7f3a", session_id = "paper-2026", job_type = "chat",
    status = "awaiting_approval", progress = 60,
    created_at = "2026-08-10T10:00:00Z", updated_at = "2026-08-10T10:01:12Z",
    metadata = list(intermediate_responses = list(
      list(type = "intermediate", content = "Reading the discussion...", sequence = 1),
      # Note the shape: `content` is a JSON string inside JSON.
      list(type = "proposed_change_batch", sequence = 2, content = jsonlite::toJSON(list(
        type = "batch_approval", batch_id = "ch_1", batch_total = 2,
        changes = list(
          list(
            change_id = "ch_1", operation = "edit", chunk_id = "c41",
            old_html = "<p>The effect was very significant.</p>",
            new_html = "<p>The effect was significant (p = 0.03).</p>",
            ai_explanation = "Replaced an intensifier with the reported statistic"
          ),
          list(
            change_id = "ch_2", operation = "delete", chunk_id = "c58",
            old_html = "<p>This clearly proves the mechanism.</p>",
            new_html = NULL,
            ai_explanation = "Removed a causal claim the data does not support"
          )
        )
      ), auto_unbox = TRUE, null = "null"))
    ))
  ),
  done = list(
    job_id = "job_7f3a", session_id = "paper-2026", job_type = "chat",
    status = "completed", progress = 100,
    created_at = "2026-08-10T10:00:00Z", updated_at = "2026-08-10T10:02:40Z",
    result = list(response = "Applied 1 change.", session_id = "paper-2026")
  )
)

json <- function(x) {
  httr2::response(
    status_code = 200, headers = list(`Content-Type` = "application/json"),
    body = charToRaw(jsonlite::toJSON(x, auto_unbox = TRUE, null = "null"))
  )
}

approved_yet <- FALSE
transport <- function(req) {
  path <- httr2::url_parse(req$url)$path
  if (grepl("documents/upload", path)) return(json(payloads$upload))
  if (grepl("chat/async", path)) return(json(payloads$started))
  if (grepl("approve", path)) {
    approved_yet <<- TRUE
    return(json(list(status = "ok")))
  }
  if (grepl("^/v1/jobs/", path)) {
    return(json(if (approved_yet) payloads$done else payloads$review))
  }
  if (grepl("documents/export", path)) {
    return(httr2::response(200L, body = charToRaw("PK\003\004 ...docx bytes...")))
  }
  stop("unrecorded call: ", path)
}
```

## The client

In real use you write this and nothing else:

```{r real-client, eval = FALSE}
client <- sd_client() # reads SUPERDOCS_API_KEY
sd_verify_key(client) # cheap check that the key works
```

Keep the key in `~/.Renviron`, never in the script. Here we point at the
recorded transport instead, and cap the client at three operations so a mistake
in this vignette could not cost anything even if it were live:

```{r client}
client <- sd_client(
  api_key = "sk_not_a_real_key",
  budget = sd_budget(max_operations = 3, max_pages = 40),
  transport = transport
)
client
```

## Upload the rendered paper

`sd_upload()` streams the file from disk rather than reading it into R, so a
long manuscript with figures does not have to fit in memory twice. Uploading
is not billable.

```{r upload}
manuscript <- tempfile(fileext = ".docx")
writeBin(charToRaw("PK... a rendered manuscript ..."), manuscript)

doc <- sd_upload(manuscript, client, session_id = "paper-2026")
doc
```

The session id is yours to choose and worth choosing deliberately: it is how
you come back to this document tomorrow, or after R crashes.

## Ask for a targeted edit

Targeted instructions land better than sweeping ones. "Improve the paper"
gives the model licence to rewrite your argument; the instruction below does
not.

```{r edit}
job <- sd_edit(
  doc,
  paste(
    "In the discussion, replace vague intensifiers with the reported statistics,",
    "and remove any causal claim the results section does not support.",
    "Do not touch the methods, the citations, or the numbers."
  )
)
job
```

`approval_mode` defaults to `"ask_every_time"`. Nothing has changed in the
document yet.

## Wait for it

An edit on a real paper takes anywhere from ten seconds to several minutes,
with nothing visible in between. That is the API working. `sd_wait()` polls
with exponential backoff so a long job is not hammered, and gives up on your
schedule rather than the server's:

```{r wait}
job <- sd_wait(job, timeout = 300, verbose = FALSE)
job$status
sd_awaiting_kind(job)
```

Two different pauses report `awaiting_approval`, which is why
`sd_awaiting_kind()` exists. `"change_review"` is answered by `sd_approve()`;
`"continue_prompt"` -- a large edit asking whether to keep going -- is answered
by `sd_continue()`. Sending the wrong one is rejected with a 409.

If `sd_wait()` times out, nothing is lost: the job keeps running server-side,
and `sd_wait(job)` picks it back up. Jobs are deleted an hour after they are
created, so that is the real deadline.

## Read what it proposes

```{r changes}
changes <- sd_changes(job, refresh = FALSE)
changes
```

It is an ordinary data frame, so you can filter it like one:

```{r changes-df}
changes[, c("change_id", "operation", "ai_explanation")]
```

### The second parse, and why you did not have to think about it

Proposed changes reach a client in two shapes. Some arrive as ordinary JSON
objects in `metadata$pending_changes`. The batched ones arrive inside
`metadata$intermediate_responses`, where the payload sits in a `content` field
that is *itself a JSON-encoded string* -- the same double encoding the SSE
`proposed_change_batch` event uses.

Parse the response once and `content` is still a character scalar; every field
you read off it comes back empty. That is the single most common reason an
integration renders a review UI where every row is blank. `sd_changes()` makes
the second `jsonlite::fromJSON()` call for you. Here is the raw payload it
started from:

```{r double-parse}
raw_event <- payloads$review$metadata$intermediate_responses[[2]]
substr(raw_event$content, 1, 60)

# One parse gets you a string. Two gets you the changes.
class(raw_event$content)
length(jsonlite::fromJSON(raw_event$content, simplifyVector = FALSE)$changes)
```

## Approve the part you agree with

This is the gate. `change_ids` has no default, and there is no
approve-everything switch, because an unreviewed edit to a paper you are about
to submit is the whole thing this package exists to prevent.

Here we take the statistic correction and reject the deletion, telling the
model why:

```{r approve}
keep <- changes$change_id[changes$operation == "edit"]
drop <- changes$change_id[changes$operation == "delete"]

job <- sd_approve(job, keep)
job <- sd_deny(job, drop, feedback = "Keep the sentence; soften it instead of cutting it.")
job <- sd_wait(job, verbose = FALSE)
job$status
```

Denying with feedback can produce a revised proposal, so a review can go
several rounds. Keep polling until the status is `completed`.

## Export

Exports are free, so export often rather than at the end.

```{r export}
out <- file.path(tempdir(), "manuscript-revised.docx")
sd_export(doc, format = "docx", path = out)
file.size(out) > 0
```

`docx`, `pdf`, `html`, `markdown` and `txt` are the available formats. If the
render hit something non-fatal -- an image that would not download, a diagram
that timed out -- `sd_export()` raises an R warning carrying the codes rather
than handing you a file that quietly lost a figure.

## The knitr shortcut

`sd_knit()` is the whole thing in one call: it renders the `.Rmd` with
`rmarkdown::render()` and then runs upload, edit, review, approve, export over
the result.

```{r knit, eval = FALSE}
result <- sd_knit(
  "paper.Rmd",
  "Tighten the discussion; keep every citation and number exactly as written.",
  output_format = "word_document",
  export_path = "paper-revised.docx"
)
result$changes
```

The review step is a function you supply. The default,
`sd_review_console()`, asks you in an interactive session and refuses to guess
in a script -- which means a scheduled job cannot approve anything by
accident. Two other useful shapes:

```{r review-fns, eval = FALSE}
# Prepare the work, decide later.
sd_knit("paper.Rmd", "...", review = sd_review_none)

# A rule you can defend: accept wording edits, never accept deletions.
sd_knit("paper.Rmd", "...", review = function(changes) {
  changes$change_id[changes$operation == "edit"]
})
```

Whatever the function returns, every proposed change gets an explicit decision
-- the ids you name are approved and the rest are denied -- so a job never
sits blocking your session waiting for a review nobody is going to give.

## Not spending more than you meant to

`sd_budget()` is attached to the client and enforced before requests leave your
machine.

```{r budget}
sd_ops_used(client)
sd_ops_remaining(client)

tiny <- sd_client(api_key = "sk_x", budget = sd_budget(max_operations = 0), transport = transport)
try(sd_edit(sd_document(tiny, "paper-2026"), "Rewrite everything."))
```

`max_pages` refuses an oversized document after the free upload and before any
billable edit. The API reports chunks, not pages, so the gate compares against
`ceiling(chunks_count / chunks_per_page)` -- an estimate, with a default of 12
chunks per page that suits a double-spaced manuscript. Tune it for your
documents:

```{r pages}
strict <- sd_client(
  api_key = "sk_x",
  budget = sd_budget(max_pages = 5, chunks_per_page = 12),
  transport = transport
)
try(sd_upload(manuscript, strict, session_id = "paper-2026"))
```

And `small_sample = TRUE` is the cheap first run: one operation, three pages,
fastest model tier.

## Picking up after a crash

Nothing about the workflow lives only in your R session. The session id and
the job id are enough to resume:

```{r resume}
doc <- sd_document(client, session_id = "paper-2026")
job <- sd_job(client, job_id = "job_7f3a", session_id = "paper-2026")

# Poll it again to find out where it got to. Here it has already finished.
sd_wait(job, verbose = FALSE)$status
```

Save both to disk as soon as you have them. A job id older than an hour is
gone rather than slow -- re-run the edit in that case.

## When something goes wrong

Errors name the cause and the fix, and carry a class so you can branch:

```{r errors, error = TRUE}
broken <- sd_client(
  api_key = "sk_x",
  transport = function(req) {
    httr2::response(
      status_code = 415,
      headers = list(`Content-Type` = "application/json"),
      body = charToRaw('{"detail": "Unsupported file type: .doc"}')
    )
  }
)
tex <- tempfile(fileext = ".doc")
writeBin(charToRaw("x"), tex)
sd_upload(tex, broken)
```

That one never reached the network: the extension check runs locally. The
classes are `sd_api_error` (plus `sd_http_<status>`), `sd_budget_error`,
`sd_timeout_error` and `sd_job_failed`.
